---
title: "Olympic Dam Mobile GIS Efficiency Gain Supplementary Materials"
author: "Ben Marwick"
date: "Monday, September 22, 2014"
output: 
  html_document:
    theme: readable
---

This document presents the statistical analysis in the 
paper 'Efficiency gains with the use of mobile 
Geographic Information Systems for intensive 
archaeological survey of Holocene forager sites in 
arid South Australia'.

First, we input the data to reproduce the table that appears
in the paper. 

```{r, get-data, results="asis"}
# read data in
dat <- read.table(text = "
Year Pre.2007  y2007	y2008	y2009
Area.surveyed.(km2)	35	95	275	145
Rate.of.survey.(km2/person/day)	0	0.3	0.3	0.3
Total.no.persondays	0 317	917	483
Approximate.number.of.sites.recorded	665	3900	6600	5500
No.sites.recorded/person/day	 3.5	12.3	7.2	11.3
Frequency.of.occurrence.of.sites/km2	19	41	24	38", row.names = 1, header = TRUE)
# print table
knitr::kable(dat)
```

Second, we compute the correlation between the rate of recording
and the density of sites.

```{r cor-rate-density}
# transpose dataframe for conviennence 
dat_t <- data.frame(t(dat))

# correlation of sites recorded by density of sites
dat_cor <- cor.test(dat_t[,5], dat_t[,6], method = 'pearson')
```

The correlation is high (r = `r unname(dat_cor$estimate)`)
and significant (p = `r unname(dat_cor$p.value)`). So we 
need to remove this effect to get a reliable estimate of
the difference made by the introduction of the mobile GIS
technology in the 2007-2009 fieldwork. 

Third, we can isolate the effect of site density on recording 
rate by computing a linear model for the 2007-2009 data 
and using this model to predicting the site recording rate 
given the density for the pre-2007 data

```{r lm-2007-2009-data}
# linear model for sites recorded by density of sites
# during  2007-2009 

# subset 2007-2009 data
dat_t_s <- dat_t[2:4,]

# linear model for sites recorded by density of sites, 2007-2009
summary(dat_t_s_lm <- lm(dat_t_s[,5] ~ dat_t_s[,6]))
```

With this model we can predict how many sites we'd expect
to be recorded with a mobile GIS given 19 sites per square 
kilometer (the site density for the pre-2007 data), which 
is `r round(19 * unname(dat_t_s_lm[[1]][2]) + unname(dat_t_s_lm[[1]][1]),2)`.

The difference of 2.2 sites/person/day between the 
actual pre-2007 average rate of 3.5 and the extrapolated 
average rate of 5.7 is most likely due to the use of mobile
GIS. This represents a `r round(2.2/3.5,2)*100` percent 
increase in the site recording rate.

#### Dependencies for this repository
```{r}
sessionInfo()
```
